\LARGE \bf A Hybrid Control Framework for Autonomous Vehicles at Uncontrolled Intersections

### 2019-01-16 conference/iv2019/manuscript.tex
chugging along through IV2019 paper

% The following packages can be found on http:\\www.ctan.org
\usepackage{graphics} % for pdf, bitmapped graphics files
\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

\title{\LARGE \bf
A Hybrid Control Framework for Navigating Pedestrian Interaction at Mid-Block Intersections}

\author{Nitin R. Kapania$^{1}$, Sarah M Thornton, Francesco Borrelli, J Christian Gerdes% <-this % stops a space
\thanks{*This work was supported by the Hyundai Center of Excellence}% <-this % stops a space
\thanks{$^{1}$Nitin Kapania is with the Dept. of Mechanical Engineering, Stanford University, and is a visiting scholar at UC Berkeley}%
}

@@

As autonomous vehicles (AVs) inch closer to reality, a central requirement for acceptance will be earning the trust of humans in everyday driving situations. In particular, the interaction between AVs and pedestrians is of
high importance, as every human is a pedestrian at some point of the day. This paper considers the interaction of a pedestrian and an autonomous vehicle at a mid-block, unsignalized intersection where there is ambiguity
over when the pedestrian should cross and when and how the vehicle should yield. By modeling pedestrian behavior through the concept of gap acceptance, the authors show that a hybrid controller with just four states allows an autonomous vehicle to successfully interact with a pedestrian across a continuous spectrum of possible crosswalk-entry behaviors. The controller is validated in simulation through comparison with a previously published control policy obtained through solution of a POMDP, and experimental results are provided on a Hyundai Genesis vehicle for a virtual pedestrian.

@@

The need to further understand pedestrian behavior for autonomous driving has created growing body of interdisciplinary research. One branch of research has focused around modeling pedestrian behavior given various sensor inputs. Keller et al \cite{Keller2014} presented a study on pedestrian path prediction and action classification (e.g. crossing vs. waiting) using Gaussian process dynamical models and trajectory matching from optical data. Several other techniques for pedestrian trajectory prediction have been proposed, including LQR \cite{Batkovic}, set-based reachability analysis \cite{Koschi2018}, and Markov processes \cite{Karasev2016}

Another branch of literature is focused on developing binary predictive models for pedestrian crossing. While typically studied for purposes of road design, this body of literature holds promising insights for AV designers. For example, Schroeder and Rouphail \cite{Schroeder2011} explored factors associated with driver yielding behavior at unsignalized pedestrian crossings. Using logistic regression, the authors found that drivers are more likely to yield to assertive pedestrians who walk briskly in their approach to a crosswalk. Kadali and Perumal \cite{RaghuramKadali2012} studied the ``gap acceptance" behavior of pedestrians at mid-block crosswalks through a video graphic survey, and found that the gap accepted for crossing was explained by factors such as crossing direction, vehicle speed, and pedestrian age. Yannis et al. \cite{Yannis2013} and \cite{Sun2002} also found that gap acceptance was influenced by the size of the oncoming vehicle and the presence of other pedestrians.  Lee and Aty \cite{Lee2005} studied interactions in the form of crashes, and found crashes were linked with higher daily traffic.

@@

\subsection{Statement of Contributions}
While there exists a wide body of research on the distinct subjects of pedestrian modeling and AV control, there is a lack of research that combines the two. In other words, significant research has been conducted to understand the likelihood of pedestrian crossing given certain traffic conditions and pedestrian demographics \cite{Schroeder2011} -\cite{Lee2005}, and a small but growing body of literature develops control strategies for pedestrian avoidance \cite{Bandyopadhyay}-\cite{Thornton2018}. However, what is missing is a contribution that explicitly considers tests whether a proposed control strategy is robust to the variety of pedestrian behaviors that have been observed from experimental studies on real roads. What is also missing is analysis of how this controller should behave across multiple traffic scenarios - for example, the navigation problem is different for the pedestrian crossing on the opposing side of traffic, and also depends on which particular lane the vehicle is in.

This paper aims to address this gap by developing a state-machine based control architecture that accounts for several distinct pedestrian modes of behavior at an unsignalized crosswalk. An unsignalized crosswalk is chosen as it is the trickiest crosswalk for a pedestrian and vehicle to navigate, as there is no explicit declaration of whose turn it is cross. Simulations show that the state machine controller is able to handle a continuous spectrum of pedestrian \textit{gap acceptance} behavior, tolerating a range of highly conservative to highly aggressive pedestrians. Moreover, this is possible with just four distinct states. Simulations are also used to compare the closed-loop controller with the controller developed by \cite{Thornton2018}. Finally, experimental results are shown on a real vehicle to demonstrate the feasibility of the proposed state machine controller.

\section{Problem Description and System Requirements}

@@

One of the major factors that determines pedestrian crossing behavior is \textit{gap acceptance}, defined in this paper as:

\begin{align}
\mathrm{gap} = \frac{\mathrm{distance\hspace{1mm}to\hspace{1mm}crosswalk}}{\mathrm{vehicle\hspace{1mm}speed}}
\end{align}

Also known as time to collision (TTC) - the safety gap is a measure of how much time there is before the vehicle would enter the crosswalk if it kept it's current speed constant. When faced with a stream of traffic at a crosswalk, a pedestrian inherently decides how much of a gap to accept before crossing. The average gap acceptance is reported in the literature to be between 3-7 seconds, meaning pedestrians usually do not cross if the vehicle would enter the crosswalk in under three seconds \cite{DiPietroCharlesMandKing1970}, and are very likely to cross when they have more than seven seconds \cite{Schmidt2009}.

\subsection{Controller Objectives}

Given the multiple stakeholders present (the autonomous vehicle and its passengers, the pedestrian(s), and the other vehicles in the traffic stream), it is important to consider the values and needs of all participants when determining the objectives of a control framework. This paper determines controller objectives based on the three-part value sensitive design (VSD) framework proposed by Thornton et al. 
\cite{Thornton2018a}

@@

@@

\section{Experimental Results}

>>>>

% The following packages can be found on http:\\www.ctan.org
\usepackage{graphics} % for pdf, bitmapped graphics files
\usepackage{epsfig} % for postscript graphics files
\usepackage{algorithm}
\usepackage{algpseudocode}
%\usepackage{align}
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

\title{\LARGE \bf
A Hybrid Control Framework for Autonomous Vehicles at Uncontrolled Intersections}

\author{Nitin R. Kapania$^{1}$, Francesco Borrelli, J Christian Gerdes% <-this % stops a space
\thanks{*This work was supported by the Hyundai Center of Excellence}% <-this % stops a space
\thanks{$^{1}$Nitin Kapania is with the Dept. of Mechanical Engineering, Stanford University, and is a visiting scholar at UC Berkeley}%
}

@@

As autonomous vehicles (AVs) inch closer to reality, a central requirement for acceptance will be earning the trust of humans in everyday driving situations. In particular, the interaction between AVs and pedestrians is of
high importance, as every human is a pedestrian at some point of the day. This paper considers the interaction of a pedestrian and an autonomous vehicle at a mid-block, uncontrolled intersection where there is ambiguity
over when the pedestrian should cross and when and how the vehicle should yield. By modeling pedestrian behavior through the concept of gap acceptance, the authors show that a hybrid controller with just four distinct modes allows an autonomous vehicle to successfully interact with a pedestrian across a continuous spectrum of possible crosswalk-entry behaviors. The controller is validated in simulation through comparison with a previously published control policy obtained through solution of a POMDP, and experimental results are provided on a Hyundai Genesis vehicle for a virtual pedestrian.

@@

The need to further understand pedestrian behavior for autonomous driving has created a growing body of interdisciplinary research. One branch of research has focused around modeling pedestrian behavior given various sensor inputs. Keller et al \cite{Keller2014} presented a study on pedestrian path prediction and action classification (e.g. crossing vs. waiting) using Gaussian process dynamical models and trajectory matching from optical data. Several other techniques for pedestrian trajectory prediction have been proposed, including LQR \cite{Batkovic}, set-based reachability analysis \cite{Koschi2018}, and Markov processes \cite{Karasev2016}

Another branch of literature is focused on developing predictive models for pedestrian crossing. While typically studied for purposes of road design, this body of literature holds promising insights for AV designers. For example, Schroeder and Rouphail \cite{Schroeder2011} explored factors associated with driver yielding behavior at unsignalized pedestrian crossings. Using logistic regression, the authors found that drivers are more likely to yield to assertive pedestrians who walk briskly in their approach to a crosswalk. Kadali and Perumal \cite{RaghuramKadali2012} studied the ``gap acceptance" behavior of pedestrians at mid-block crosswalks through a video graphic survey, and found that the gap accepted for crossing was explained by factors such as crossing direction, vehicle speed, and pedestrian age. Yannis et al. \cite{Yannis2013} and \cite{Sun2002} also found that gap acceptance was influenced by the size of the oncoming vehicle and the presence of other pedestrians.  Lee and Aty \cite{Lee2005} studied interactions in the form of crashes, and found crashes were linked with higher daily traffic.

@@

\subsection{Statement of Contributions}
Significant research has been conducted to understand the likelihood of pedestrian crossing given certain traffic conditions and pedestrian demographics \cite{Schroeder2011} -\cite{Lee2005}, and a small but growing body of literature develops control strategies for pedestrian avoidance \cite{Bandyopadhyay}-\cite{Thornton2018}. However, what is missing is a contribution that explicitly  tests whether a proposed control strategy is robust to the variety of pedestrian behaviors that have been observed from experimental studies on real roads. What is also missing is analysis of how this controller should behave across multiple traffic scenarios - for example, the navigation problem is different for the pedestrian crossing on the opposing side of traffic, and also depends on which particular lane the vehicle is in.

This paper aims to address this gap by developing a hybrid control architecture that accounts for several distinct pedestrian modes of behavior at an unsignalized crosswalk. An unsignalized crosswalk is chosen as it is the trickiest crosswalk for a pedestrian and vehicle to navigate, as there is no explicit declaration of whose turn it is cross. Simulations show that the state machine controller is able to handle a continuous spectrum of pedestrian \textit{gap acceptance} behavior, tolerating a range of highly conservative to highly aggressive pedestrians. Moreover, this is possible with just four distinct states. Simulations are also used to compare the closed-loop controller with the controller developed by \cite{Thornton2018}. Finally, experimental results are shown on a real vehicle to demonstrate the feasibility of the proposed state machine controller.

\section{Problem Overview}

@@

One of the major factors that determines pedestrian crossing behavior at uncontrolled intersections is \textit{gap acceptance}, defined in this paper as:

\begin{equation}
\mathrm{gap} = \frac{\mathrm{distance\hspace{1mm}to\hspace{1mm}crosswalk}}{\mathrm{vehicle\hspace{1mm}speed}}
\end{equation}

Similar to time to collision (TTC), the safety gap is a measure of how much time there is before the vehicle would enter the crosswalk if it kept it's current speed constant. When faced with a stream of traffic at a crosswalk, a pedestrian inherently decides how much of a gap to accept before crossing. The average gap acceptance is reported in the literature to be between 3-7 seconds, meaning pedestrians usually do not cross if the vehicle would enter the crosswalk in under three seconds \cite{DiPietroCharlesMandKing1970}, and are very likely to cross when they have more than seven seconds \cite{Schmidt2009}.

\subsection{Problem Formalization and System Requirements}

A diagram of the relevant state variables is shown below in Fig.~\ref{fig:diagram1}. We assume that knowledge of distance $d$ to a fixed stopping point 3-5 meters ahead of the crosswalk can be estimated by a perception system, and that the velocity of the vehicle $\dot{d}$ is known as well. Furthermore, we also assume the perception system is able to determine the position of the pedestrian $x_p$ within the crosswalk, and is able to determine a velocity estimate $\dot{x_p}$ for the pedestrian as well.

\begin{figure}
\centering
\includegraphics[width=1in]{figures/diagram.eps}
\caption{Description of relevant problem variables.}
\label{fig:diagram1}
\end{figure}

The relevant control problem is therefore to determine an appropriate longitudinal acceleration command for the vehicle, $u = \ddot{d}$, as a closed-loop function of the vehicle and pedestrian states:

\begin{equation}
\ddot{d} = f(d, \dot{d}, x_p, \dot{x_p})
\end{equation}

In general, this control problem is difficult as there are multiple stakeholders and therefore multiple competing objectives that must be considered. For example, relevant stakeholders for this problem are the autonomous vehicle and its passengers, the pedestrian(s), and the other vehicles in the traffic stream.

As a result, an algorithm that simply yields every time a pedestrian is near the crosswalk is likely to be too conservative, as there is a likelihood the pedestrian is waiting for a larger gap in the traffic stream to cross or is walking through the sidewalk without intending to cross in the first place. To formalize the system requirements for this multi-stake holder problem, we propose the following \textit{design constraints}, extended from the engineering specifications proposed by Thornton \cite{Thornton2018}.

\begin{center}
\begin{tabular}{ l| l }
\label{tb:specs}
 \textbf{Engineering} & \textbf{Design} \\
 \textbf{Specification} & \textbf{Constraint}\\\hline 
 1. Safety and Legality & a. Avoid collisions for all\\
                     & possible accepted gaps                        \\
                     &b. Ability to follow stop/yield    \\
                     &laws for all 50 U.S. states\\

&\\
 2. Efficiency & a. Average speed: Minimize  \\ 
               & deviation from speed of traffic \\    
               & b. Wait for pedestrian crossing \\
               & event to materialize before yielding. \\

               & \\
 3. Smoothness & a. Low accel: Nominal brake / \\ 
 			   & accel limits of 2 m/s/s \\
 			   & b. Stop within 3-5 meters\\
 			   & of crosswalk

\end{tabular}
\end{center}

@@

The proposed control architecture is shown below in Fig.~\ref{fig:hybridController}, and the algorithm for operation is shown in Algorithm \ref{alg:hybridController}. At heart, the control algorithm uses a feedback-feedforward methodology to compute desired acceleration commands $\ddot{d}$, but computes these commands differently depending on which of four discrete states the controller is in. The next sections will describe the four states.

\begin{figure}
\centering
\includegraphics[width=3.5in]{figures/stateDiagram.eps}
\caption{Schematic of hybrid control structure.}
\label{fig:hybridController}
\end{figure}

\begin{algorithm} % enter the algorithm environment
\caption{Calculate $\ddot{d} = f(d, \dot{d}, x_p, \dot{x}_p)$} % give the algorithm a caption
\label{alg:hybridController} % and a label for \ref{} commands later in the document
\begin{algorithmic}
\State $state \leftarrow DRIVING$
\While {$\mathbf{true}$}
\If {$state = DRIVING$}
	\State $\ddot{d} \gets k_s(\dot{d}-v_\mathrm{speed limit})$ 
	\If{$d > 0 \hspace{1mm} \mathbf{and} \hspace{1mm} \mathrm{inCrosswalk}(x_p)$}
		\If{$\mathrm{timeAdvantage}(x_p, \dot{x_p}, d, \dot{d}) > t_{max}$} 
		\State $state \leftarrow DRIVING$

		\ElsIf{$d > \frac{\dot{d}^2}{2a_\mathrm{cmfrt}}$}
		\State $state \leftarrow YIELDING$

	\ElsIf{$d < \frac{\dot{d}^2}{2a_\mathrm{cmfrt}} \hspace{1mm} \mathbf{and} \hspace{1mm} d > \frac{\dot{d}^2}{2a_\mathrm{max}}$}
	\State $state \leftarrow HARD\hspace{1mm}BRAKING$

	\Else
	\State $state \leftarrow SPEED\hspace{1mm}UP$

	\EndIf
	\EndIf
	\EndIf
\vspace{3mm}
\If {$state = YIELDING$}
\If {$d > \frac{\dot{d}^2}{2a_\mathrm{cmfrt}} + t_{delay}\dot{d}$}
	\State $\ddot{d}_{des} = 0$
	\State $\dot{d}_{des} = v_\mathrm{speed limit}$
\Else 
	\State $\ddot{d}_{des} = -a_\mathrm{cmf}$
	\State $\dot{d}_{des} = \mathrm{getDesiredSpeed}(d, \dot{d})$
\EndIf 
	\State $\ddot{d} = \ddot{d}_{des} + k_s(\dot{d} - \dot{d}_{des})$

\If {$\textbf{not} \hspace{1mm} \mathrm{inCrosswalk}(x_p)$} \State $state \leftarrow DRIVING$

\EndIf
\EndIf

\vspace{3mm}

\If {$state = HARD\hspace{1 mm}BRAKING$}
	\State $\dot{d}_{des} = \mathrm{getDesiredSpeed}(d, \dot{d})$
	\State $\ddot{d} = -\frac{\dot{d}^2}{2d} + k_s(\dot{d} - \dot{d}_{des})$ 	
	\If {$\textbf{not} \hspace{1 mm} \mathrm{inCrosswalk}(x_p)$} \State $state \leftarrow DRIVING$

	\EndIf
	\EndIf

\EndIf
\If {$state = SPEED\hspace{1mm}UP$}
	\State $\ddot{d} = a_\mathrm{cmf}$ 	
	\If {$\textbf{not} \hspace{1mm} \mathrm{inCrosswalk}(x_p)$} \State $state \leftarrow DRIVING$

	\EndIf
	\EndIf

\EndWhile

\end{algorithmic}
\end{algorithm}

\subsection{Nominal Driving}
In the nominal state, the vehicle attempts to drive through the crosswalk at the speed limit $v_\mathrm{speedlimit}$. A simple proportional speed control is applied to keep the vehicle at the speed limit, achieving design constraint 2.a.:

\begin{equation}
	\ddot{d} = k_s(\dot{d} - v_\mathrm{speedlimit})
\end{equation}

The controller remains in the driving mode unless the pedestrian begins to enter the crosswalk while $d > 0$. To allow more time for the controller to respond, we allow the $\mathrm{inCrosswalk}(x_p, \dot{x}_p)$ indicator function to be triggered the moment the pedestrian approaches the crosswalk from the sidewalk - e.g. once the pedestrian velocity $\dot{x}_p$ becomes non-zero.

 Note that according to design constraint 1.b, the definition of entering the crosswalk will vary according to which of the fifty US states the vehicle is in. For example, in California, the pedestrian entering any portion of the crosswalk should trigger the control mode to change - while in Louisiana, the pedestrian entering the same half of the roadway as the vehicle should trigger the change \cite{NCSL2018}.

Once the pedestrian enters the crosswalk, the algorithm decides which mode to enter next.

\subsubsection{Time Advantage}

Frequently he vehicle will pass through the intersection well before the pedestrian will reach the vehicle's lane. Assuming the vehicle is operating in a US state where stopping is not explicitly required, design constraint 2.a suggests it is desirable for the vehicle to continue through the intersection.

This can be formalized by calculating the \textit{time advantage} $t_adv$, also known as the \textit{time to collision} as follows:

\begin{equation}
t_{adv} = \frac{d}{\dot{d}} - \frac{x_v - x_p}{\dot{x_p}}
\end{equation}

Where $x_v$ is the $x$ position of the vehicle in the crosswalk. The control algorithm in this work will explicitly allow the vehicle to continue driving through the intersection as long as $t_{adv}$ exceeds a specfiied threshold $t_{max}$.

\subsubsection{Yielding}

If the time advantage is not sufficient for the vehicle to pass through, the autonomous vehicle must slow down. The preferred option to meet design constraint 3.a is to brake at a low, comfortable deceleration $-a_{cmf}$ for the passenger. Note that design constraint 3.b is satisfied by having $d=0$ be defined to be a point several meters ahead of the crosswalk.

Comfortable braking at a deceleration $-a_{cmf}$ is possible if sufficient braking distance exists. The braking distance is derived kinematically as

\begin{equation}
d_{brake} = \frac{\dot{d}^2}{2a_{cmf}}
\end{equation}

If this holds true when the pedestrian starts to enter the crosswalk, the car enters a $Yielding$ state.

\subsubsection{Hard Braking}

In the unlikely event a pedestrian begins to cross when the time gap is low (e.g. under 3 seconds), the vehicle will need to prioritize design constraint 1.a and come to a stop at a deceleration higher than $-a_{cmf}$. This occurs when the following conditions holds:

\begin{equation}
\frac{\dot{d}}{2a_{max}} < d < \frac{\dot{d}^2}{2a_{cmf}}
\end{equation}

Where $a_{max}$ is the largest deceleration magnitude allowed by the tire-road friction.

\subsection{Speeding Up}

As a final condition, consider the case where the pedestrian enters the crosswalk just as the vehicle is crossing $d = 0$ - in this case, it makes little sense to decelerate, as the vehicle has insufficient space to stop before the crosswalk and will risk being rear-ended (Design Constraint 2.a) or stopping in the crosswalk. It makes more sense in this condition for the vehicle to speed up and exit the crosswalk quickly. As a result, if $d < \frac{\dot{d}}{2a_{max}}$, the vehicle speeds up.

\subsection{Yielding}

In the yielding state, the vehicle follows a different version of the feedback-feedforward dynamics. The vehicle first follows the speed limit until it reaches the critical value of $\frac{\dot{d}^2}{2a_{cmf}}$. In practice, an additional term $t_delay\dot{d}$ is added to compensate for the delay in brake communication.

At the critical istance, the vehicle decelerates at the uniform yielding deceleration of $-a_{cmf}$, with an additional feedback term $k_s(\dot{d} - \dot{d}_des$), where $\dot{d}_des$ is the desired velocity. The feedback term helps bring the vehicle speed to 0 at the desired stopping point.

The desired velocity profile $\dot{d}(d)$ for a vehicle decelerating at constant acceleration is given via integration:

\begin{equation}
\dot{d}(d) = \sqrt{2a_{cmf}(d-d_o) + \dot{d}_o^2}
\end{equation}

Where $d_o$ and $\dot{d}_o$ is the value of $d$ and $\dot{d}$ when the controller first enters the yielding state. Simple inspection shows that the desired velocity is $\dot{d}_o$ when $d = d_o$ and is 0 when $d = 0$ (recall that we enter the yielding state when $d_o = \frac{\dot{d_o}^2}{2a_{cmf}}$.

The controller exits the yielding mode and returns to the driving mode once the pedestrian clears the crosswalk.

\subsection{Hard Braking}

In the hard braking state, the vehicle follows a similar feedback-feedforward algorithm, but the desired feedforward deceleration in this case is given by $\ddot{d} = -\frac{d^2}{2d}$. Integrating to find the desired speed profile $\dot{d}(d)$ yields the following deceleration:

\begin{equation}
\dot{d}(d) = \frac{\dot{d}_o}{\sqrt{d_o}}\sqrt{d}
\end{equation}

Where $d_o$ and $\dot{d}_o$ is the value of $d$ and $\dot{d}$ when the controller first enters the braking state. Again, the controller exits the braking mode and returns to the driving mode once the pedestrian clears the crosswalk.

\subsection{Speed Up}

In the speed up state, there is no need to follow an exact speed profile as the vehicle is merely trying to exit the crosswalk area slightly faster. In this case, the commanded acceleration $\ddot{d} = a_{cmf}$ until the pedestrian clears the crosswalk.

@@

\subsection{Evaluation Methodology}

In order to check performance of the algorithm, both simulation and experimental studies were undertaken. The simulation results focused on testing Algorithm 1 against the design requirements from Table \ref{tb:specs} over many pedestrian crossings.

The algorithm is validated in simulation via a built-from-scratch crosswalk simulation environment (Fig. \ref{fig:simFramework}) developed in Python. The simulation environment is built on the ROS \cite{ROS} robotics interface to enable easy porting of the control architecture to the experimental vehicle in Section \ref{sec:expres}. The simulation includes features such as multiple lanes, the pedestrian crossing on the right or the left, and animation capabilities.

\begin{figure}
\centering
\includegraphics[width=3.5in]{figures/simFramework.png}
\caption{Crosswalk simulation environment for algorithm validation.}
\label{fig:simFramework}
\end{figure}

To simulate random pedestrian crossing events, a pedestrian is spawned in an area near the crosswalk. The pedestrian accepts a randomly sized time gap in the traffic stream to cross the road at constant speed $\dot{x}_p$. The size of the time gap is drawn from a normal distribution with mean $\mu_{gap}$ and standard deviation $\sigma_{gap}$. Statistics for $\mu_{gap}$ and $\sigma_{gap}$ are obtained from a video graphic analysis by Feliciani et al. \cite{Feliciani2017}.

The ego vehicle's response to the pedestrian is simulated over Algorithm 1 for the parameters shown in Table \ref{tb:params}.

\begin{table}[h]
\footnotesize
\begin{center}
\caption{Simulation and Controller Parameters}\label{tb:params}
\begin{tabular}{lccc}
Parameter & Symbol & Value & Units \\\hline\hline
Number of lanes & $n$ & 4 & - \\
Gap variance & $\sigma_{gap}$ & 2.5 & sec\\
Mean accepted gap & $\mu_{gap}$ & 4.0 & sec\\\hline
Speed feedback gain & $k_s$ & 2.0 & 1 / sec\\
Brake time delay    & $t_\mathrm{delay}$ & 0 & sec \\ 
Speed limit & $v_\mathrm{speedlimit}$ & 4.5 & m/s \\
Comfort acceleration & $a_\mathrm{cmf}$ & 2 & $\mathrm{m/s^2}$ \\
Maximum acceleration & $a_\mathrm{max}$ & 9 & $\mathrm{m/s^2}$ \\\hline
\end{tabular}
\end{center}
\end{table}

\subsection{Simulation Results}

Figure \ref{fig:velPlot1} shows simulation results for the case where the pedestrian enters the crosswalk on the same side of the road as the vehicle. The left side of the figure is for the case where the vehicle is on the right-most lane of a four way road, while the right side of the figure is for the case where the vehicle is in the second to right lane.

\begin{figure}
\centering
\includegraphics[width=3.5in]{figures/velPlot.eps}
\caption{Simulation results for case where pedestrian enters the crosswalk on the same side of the road as the vehicle.}
\label{fig:velPlot1}
\end{figure}

Figure \ref{fig:velPlot1} shows simulation results for the case where the pedestrian enters the crosswalk on the same side of the road as the vehicle. The left side of the figure is for the case where the vehicle is on the right-most lane of a four way road, while the right side of the figure is for the case where the vehicle is in the second to right lane.

\section{Experimental Results}
\label{sec:expres}
